Title

### 2024-05-13 paper/main.tex
Updates

\begin{abstract}
	The initial step in a causal analysis involves determining the causal
	structure between the variables from data typically in the form of a
	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
	This procedure, known as \emph{causal discovery}, has been extensively
	studied for both DAGs and SEMs. For DAGs, numerous automated algorithms
	have been developed that can automatically estimate the structure from
	a given dataset. However, the adaption of these algorithms in applied
	fields remain limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset. As a
	result, in applied disciplines, DAGs are predominantly constructed
	manually based on domain expertise which makes it vital to test the fit
	of these models to the data and potentially modifying them to improve
	the fit. On the other hand, in the field of SEMs, models are
	typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{specification search} where expert knowledge
	combined with methods like \emph{modification indices} can be used to 
	guide the modification process. In the case of DAGs, although
	statistical tests are available that can assess the global fit of
	model, but there are no methods that can guide our modification
	process. In this paper, we present a simple modification process aimed
	at helping researchers in manually constructing or modifying the DAGs
	outputted by an algorithm to improve its fit. This modification process
	identifies pairs of variables whose association is not fully explained
	by the model. Once such an unexplained association is found, an expert
	can start by adding edges between the worst offending pair of variables 
	while using their expertise to decide the direction between
	the variables. For the measure of association, in the case of
	continuous and discrete variables we can use effect sizes, however no
	such measure of association is present for mixed data. We propose a
	novel (conditional) measure of association for mixed data based on
	canonical correlations. We also present a graphical web-tool that can
	assist researchers in iteratively modifying and constructing their
	DAGs. We theoretically show that in the presence of an oracle, this
	iterative modification process is consistent in recovering the true
	DAG. Emprically, we show that this process performs comparably to PC
	and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

>>>>

\begin{abstract}
	The initial step in a causal analysis involves determining the causal
	structure between the variables from data typically in the form of a
	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
	This procedure, known as \emph{causal discovery}, has been extensively
	studied for both DAGs and SEMs. While numerous automated algorithms
	exist for estimating DAGs from datasets, their application in applied
	remain limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset. As a
	result, in applied disciplines, DAGs are predominantly constructed
	manually based on domain expertise making it vital to test the fit of
	these models to the data and potentially modifying them to improve the
	fit. Conversely, SEMs are typically constructed manually where
	researchers come up with an initial model based on domain knowledge and
	then iteratively assess its fit and modify the model according to that.
	This process is commonly known as \emph{specification search} where
	expert knowledge combined with methods like \emph{modification indices}
	is used to guide the modification process. In the case of DAGs,
	although statistical tests are available that can assess the fit of the
	model, but there are no methods that can guide our modification
	process.

	In this paper, we present a method to guide the modification process aimed
	at helping researchers in manually constructing or modifying the DAGs
	outputted by an algorithm to improve its fit. This modification process
	identifies pairs of variables whose association in data is not fully explained
	by the model. Once such an unexplained association is found, an expert
	can iteratively add edges between the worst offending pair of variables 
	while using their expertise to decide the direction between
	the variables. For the measure of association, in the case of
	continuous and discrete variables we can use effect sizes, however no
	such measure of association is present for mixed data. We propose a
	novel (conditional) measure of association for mixed data based on
	canonical correlations. We also present a graphical web-tool that can
	assist researchers in iteratively modifying and constructing their
	DAGs. We theoretically show that in the presence of an oracle, this
	iterative modification process is consistent in recovering the true
	DAG. Empirically, we show that this process performs comparably to PC
	and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}
